---
title: "CS5200 Fall 2020: Practicum 2"
author: "Chandra Davis, Evan Douglass"
output:
  pdf_document: default
---

## Overview

We've decided to work with SQLite for this practicum. As such, to work with these files you will need SQLite installed on your machine. The data we are using is provided by IMDB at https://datasets.imdbws.com/

Download all the compressed tsv files into a folder called `data/`. You can leave them in their compressed format, as that is how we will read them in to this notebook.

## Setup

Before working with the data we need to create a place to store it. This section will set up any constants needed later and connect to a SQLite database.

```{r}
# Setup SQLite database
library(RSQLite)
DB_NAME <- "imdb-data.db"
conn <- dbConnect(RSQLite::SQLite(), DB_NAME)
```

!!!Need to download the files from the web before completing the next chunk!!!
Files were manually downloaded and saved to the data folder for testing

```{r}
# Create dataframes for the IMDB files
df.akas <- read.table("data/title.akas.tsv.gz", sep='\t', fill=TRUE, header=TRUE, encoding="UTF-8")
df.nBasics <- read.table("data/name.basics.tsv.gz", sep='\t', fill=TRUE, header=TRUE, encoding="UTF-8")
df.tBasics <- read.table("data/title.basics.tsv.gz", sep='\t', fill=TRUE, header=TRUE, encoding="UTF-8")
df.crew <- read.table("data/title.crew.tsv.gz", sep='\t', fill=TRUE, header=TRUE, encoding="UTF-8")
df.ep <- read.table("data/title.episode.tsv.gz", sep='\t', fill=TRUE, header=TRUE, encoding="UTF-8")
df.prin <- read.table("data/title.principals.tsv.gz", sep='\t', fill=TRUE, header=TRUE, encoding="UTF-8")
df.ratings <- read.table("data/title.ratings.tsv.gz", sep='\t', fill=TRUE, header=TRUE, encoding="UTF-8")
```

