---
title: 'CS5200 Fall 2020: Practicum 2'
author: "Chandra Davis, Evan Douglass"
output:
  word_document: default
  pdf_document: default
  html_document:
    df_print: paged
---

## Overview

We've decided to work with SQLite for this practicum. As such, to work with these files you will need SQLite installed on your machine. The data we are using is provided by IMDB at https://datasets.imdbws.com/

## Setup

The first thing we need to do is download and process the data files we need for the practicum. These will be downloaded from IMDB and processed into smaller sample files for ease of working with in the rest of the notebook. In normal circumstances we would do this with a faster language that is better suited for file manipulation in a separate script. For the purposes of this assignment we have included the entire script below as a single chunk. If you are running this yourself and want to either download fresh copies of the data from IMDB, or process existing copies, then you should change the `download` and/or `process` variables to `TRUE` below. But be warned this will take a VERY long time depending on your computer. We have included the processed files with our submission so that you do not have to wait for this unless you want to.

```{r, results='hide', message=FALSE}
library(data.table)
library(hash)
library(readr)
library(tidyr)
library(dplyr)
library(tibble)
library(stringr)

# False by default. Change to true if you need these files or want new versions.
# This may necessitate filtering the datasets as well.
# Caution: This will take a long time and may overwhelm your computer!!! 
download <- FALSE
process <- FALSE
create <- TRUE
load <- create

# File constants
dataDir <- "./data/"
sampleDataDir <- "./sample-data/"
url <- "https://datasets.imdbws.com/"
nameBasics <- "name.basics"
titleBasics <- "title.basics"
titleAkas <- "title.akas"
titleCrew <- "title.crew"
titleEpisode <- "title.episode"
titlePrincipals <- "title.principals"
titleRatings <- "title.ratings"
gz <- ".tsv.gz"
sampleTsv <- ".sample.tsv"

# ========== Functions ==========

downloadData <- function() {
  #' Downloads all needed data files from IMDB.
  filenames <- c(nameBasics, titleBasics, titleAkas, titleCrew,
                 titleEpisode, titlePrincipals, titleRatings)
  for (name in filenames) {
    website <- paste(url, name, gz, sep="")
    destination <- paste(dataDir, name, gz, sep="")
    download.file(website, destination)
  }
}

createDataframe <- function(filename, rows=-1) {
  #' Creates a dataframe out of a tsv file from IMDB. Can specify a row limit
  #' with rows, but defaults to the whole file.
  fread(filename, na.strings = "\\N", encoding = "UTF-8", data.table = FALSE, 
        showProgress = FALSE, nrows = rows)  # fread is a faster alternative to read.tsv
}

createTitleIdSet <- function(episodeRows, nameRows) {
  #' Creates a set of title ids from the tconst values in the episode and 
  #' name tsv files. These are used due to create a representative dataset
  #' that has both episodes and actors that appear in more than one included
  #' movie (that is not a show).
  epPaths <- getFullAndSamplePaths(titleEpisode)
  namePaths <- getFullAndSamplePaths(nameBasics)
  
  df.eps <- createDataframe(epPaths$full, episodeRows)
  df.names <- createDataframe(namePaths$full, nameRows)
  
  # Separate knownForTitles
  kfTitles <- drop_na(df.names["knownForTitles"]) %>%
    mutate(knownForTitles = strsplit(as.character(knownForTitles), ",")) %>%
    unnest(knownForTitles)
  
  # Combine all titles into a set
  allTitles <- unique(
    union_all(
      as.character(df.eps$tconst),
      as.character(df.eps$parentTconst),
      kfTitles$knownForTitles
    )
  )
  titleSet = hash(allTitles, 1:length(allTitles))  # values don't matter for the set
  
  titleSet
}

createNamesIdSet <- function(princNames, dirNames, writeNames) {
  # Gather all unique name IDs into single series
  allNames <- unique(
    union_all(
      princNames,
      dirNames,
      writeNames
    )
  )
  # Add them to a hash
  nameSet = hash(allNames, 1:length(allNames))  # values don't matter for the set
  
  nameSet
}

writeEpisodeSample <- function(numRows) {
  #' Writes sample data from the episodes tsv. This can be processed faster
  #' than others because we simply select a number of rows to included from
  #' the top.
  paths <- getOldAndNewPaths(titleEpisode)
  df <- createDataframe(paths$full, numRows)
  write_tsv(df, paths$sample)
}

writeFilteredTsv <- function(filename, filterSet, colname = "tconst") {
  #' Writes an existing dataframe after filtering it on tconst
  paths <- getFullAndSamplePaths(filename)
  df <- createDataframe(paths$full)
  
  # This may take several minutes...
  if (colname == "tconst") {
    df <- filter(df, has.key(tconst, filterSet))  
  } else if (colname == "titleId") {
    df <- filter(df, has.key(titleId, filterSet))
  } else if (colname == "nconst") {
    df <- filter(df, has.key(nconst, filterSet)) 
  } # Don't filter if colname different
  
  # Write out the data
  write_tsv(df, paths$sample)
  
  # Return the filtered dataframe in case it needs further processing
  df
}

flattenAndWriteCrewToDirectors <- function(crewDataframe) {
  #' Takes the crew data and flattens the directors column so all values are
  #' on their own line and then writes it to disk. Will be used later.
  # Separate directors
  df.directors <- crewDataframe[c("tconst", "directors")] %>%
    mutate(directors = strsplit(as.character(directors), ",")) %>%
    unnest(directors)
  
  # Remove titles with no director listed
  df.directors <- drop_na(df.directors)
  
  # Write
  outfile <- paste(sampleDataDir, "title.directors", sampleTsv, sep="")
  write_tsv(df.directors, outfile)
  
  df.directors
}

flattenAndWriteCrewToWriters <- function(crewDataframe) {
  #' Takes the crew data and flattens the writers column so all values are
  #' on their own line and then writes it to disk. Will be used later.
  # Separate writers
  df.writers <- crewDataframe[c("tconst", "writers")] %>%
    mutate(writers = strsplit(as.character(writers), ",")) %>%
    unnest(writers)
  
  # Remove titles with no director listed
  df.writers <- drop_na(df.writers)
  
  # Write
  outfile <- paste(sampleDataDir, "title.writers", sampleTsv, sep="")
  write_tsv(df.writers, outfile)
  
  df.writers
}

getFullAndSamplePaths <- function(filename) {
  #' Returns the path to the original data file and the sample data file in
  #' a list.
  full <- paste(dataDir, filename, gz, sep="")
  sample <- paste(sampleDataDir, filename, sampleTsv, sep="")
  list("full" = full, "sample" = sample)
}

# ========== Actual processing ==========

if (download) {
  if (!dir.exists(dataDir)) {
    dir.create(dataDir)
  }
  downloadData()
}

if (process) {
  if (!dir.exists(sampleDataDir)) {
    dir.create(sampleDataDir)  
  }
  # Constants for selecting titles to include
  numEpRows <- 500
  numNameRows <- 50
  
  # Get titles to include
  titleIdSet <- createTitleIdSet(numEpRows, numNameRows)
  
  # Process files that we do not need to extract name ids from
  writeEpisodeSample(numEpRows)
  writeFilteredTsv(titleBasics, titleIdSet)
  writeFilteredTsv(titleAkas, titleIdSet, "titleId")
  writeFilteredTsv(titleRatings, titleIdSet)
  
  # Process files that do need to extract name ids
  df <- writeFilteredTsv(titlePrincipals, titleIdSet)
  princ.names <- as.character(df$nconst)
  df <- writeFilteredTsv(titleCrew, titleIdSet)
  dirs.names <- flattenAndWriteCrewToDirectors(df)
  write.names <- flattenAndWriteCrewToWriters(df)

  nameIdSet <- createNamesIdSet(
    princ.names,
    dirs.names$directors,
    write.names$writers
  )

  # Finally, process the names table
  writeFilteredTsv(nameBasics, nameIdSet, "nconst")
}
```

Before working with the data we need to create a place to store it. This section will set up a SQLite database and any constants needed later.

```{r}
library(RSQLite)

# Setup SQLite database
DB_NAME <- "imdb-data.db"
conn <- dbConnect(RSQLite::SQLite(), DB_NAME)
```

## Database Schema

The following chunk creates all the necessary tables in our database based on the schema presented earlier.

```{r, results='hide'}
if (create) {
  # First we need to remove any existing data,
  # for example, if this has been run before.
  drop_table <- function(table_name) {
    paste("DROP TABLE IF EXISTS ", table_name, ";", sep="")
  }
  
  dbExecute(conn,"DROP VIEW IF EXISTS actors;")
  
  # Saving rows affected to a var prevents output
  dbExecute(conn, "PRAGMA foreign_keys = OFF;") # Avoid FK checks
  curr_tables <- dbListTables(conn)
  for (table in curr_tables) {
    dbExecute(conn, drop_table(table))
  }
  dbExecute(conn, "PRAGMA foreign_keys = ON;")
  
  # Now we can create the tables
  build_table <- function(table_def) {
    CREATE <- "CREATE TABLE IF NOT EXISTS"
    paste(CREATE, table_def)
  }
  
  # Build table definition list
  tables <- c(
    build_table(
      "Formats (
        format_id INTEGER PRIMARY KEY,
        format TEXT NOT NULL
      );"
    ),
    build_table(
      "Media (
        tconst TEXT PRIMARY KEY,
        format_id INTEGER NOT NULL,
        primaryTitle TEXT,
        originalTitle TEXT,
        isAdult INTEGER, -- 0=false, else true
        startYear INTEGER,
        endYear INTEGER,
        runtimeMinutes INTEGER,
        FOREIGN KEY (format_id) REFERENCES Formats(format_id)
      );"
    ),
    build_table(
      "Ratings (
        tconst TEXT PRIMARY KEY,
        averageRating REAL,
        numVotes INTEGER,
        FOREIGN KEY (tconst) REFERENCES Media(tconst)
      );"
    ),
    build_table(
      "Episodes (
        tconst TEXT PRIMARY KEY,
        parentTconst TEXT NOT NULL,
        seasonNumber INTEGER,
        episodeNumber INTEGER,
        FOREIGN KEY (tconst) REFERENCES Media(tconst),
        FOREIGN KEY (parentTconst) REFERENCES Media(tconst)
      );"
    ),
    build_table(
      "Genres (
        genre_id INTEGER PRIMARY KEY,
        genre TEXT NOT NULL
      );"
    ),
    build_table(
      "Media_Genres (
        mg_id INTEGER PRIMARY KEY,
        tconst TEXT NOT NULL,
        genre_id INTEGER NOT NULL,
        FOREIGN KEY (tconst) REFERENCES Media(tconst),
        FOREIGN KEY (genre_id) REFERENCES Genres(genre_id),
        CONSTRAINT unique_mg_tc_gid UNIQUE (tconst, genre_id)
      );"
    ),
    build_table(
      "People (
        nconst TEXT PRIMARY KEY,
        primaryName TEXT,
        birthYear INTEGER,
        deathYear INTEGER,
        age INTEGER,
        numMovies INTEGER
      );"
    ),
    build_table(
      "Known_For_Titles (
        kt_id INTEGER PRIMARY KEY,
        nconst TEXT NOT NULL,
        tconst TEXT NOT NULL,
        FOREIGN KEY (tconst) REFERENCES Media(tconst),
        FOREIGN KEY (nconst) REFERENCES People(nconst),
        CONSTRAINT unique_kft_nc_tc UNIQUE (nconst, tconst)
      );"
    ),
    build_table(
      "Professions (
        prof_id INTEGER PRIMARY KEY,
        prof_title TEXT NOT NULL
      );"
    ),
    build_table(
      "Primary_Profession (
        pp_id INTEGER PRIMARY KEY,
        nconst TEXT NOT NULL,
        prof_id INTEGER NOT NULL,
        FOREIGN KEY (nconst) REFERENCES People(nconst),
        FOREIGN KEY (prof_id) REFERENCES Professions(prof_id),
        CONSTRAINT unique_pp_nc_pid UNIQUE (nconst, prof_id)
      );"
    ),
    build_table(
      "Categories (
        category_id INTEGER PRIMARY KEY,
        category TEXT NOT NULL
      );"
    ),
    build_table(
      "Crew (
        tconst TEXT NOT NULL,
        ordering INTEGER NOT NULL,
        nconst TEXT NOT NULL,
        category_id INTEGER NOT NULL,
        job TEXT,
        characters TEXT,
        PRIMARY KEY (tconst, ordering),
        FOREIGN KEY (tconst) REFERENCES Media(tconst),
        FOREIGN KEY (nconst) REFERENCES People(nconst),
        FOREIGN KEY (category_id) REFERENCES Categories(category_id)
      );"
    ),
    build_table(
      "Characters (
        character_id INTEGER PRIMARY KEY,
        character TEXT NOT NULL
      );"
    ),
    build_table(
      "Cast (
        cast_id INTEGER PRIMARY KEY,
        tconst TEXT NOT NULL,
        ordering INTEGER NOT NULL,
        character_id INTEGER NOT NULL,
        -- Media table PK tconst has ref. integrity through Crew Table
        FOREIGN KEY (tconst, ordering) REFERENCES Crew(tconst, ordering),
        FOREIGN KEY (character_id) REFERENCES Characters(character_id),
        CONSTRAINT unique_crew_tc_ord_tid UNIQUE (tconst, ordering, character_id)
      );"
    ),
    build_table(
      "Media_Directors (
        md_id INTEGER PRIMARY KEY,
        tconst TEXT NOT NULL,
        nconst TEXT NOT NULL,
        FOREIGN KEY (tconst) REFERENCES Media(tconst),
        FOREIGN KEY (nconst) REFERENCES People(nconst),
        CONSTRAINT unique_md_nc_tc UNIQUE (nconst, tconst)
      );"
    ),
    build_table(
      "Media_Writers (
        mw_id INTEGER PRIMARY KEY,
        tconst TEXT NOT NULL,
        nconst TEXT NOT NULL,
        FOREIGN KEY (tconst) REFERENCES Media(tconst),
        FOREIGN KEY (nconst) REFERENCES People(nconst),
        CONSTRAINT unique_mw_nc_tc UNIQUE (nconst, tconst)
      );"
    ),
    build_table(
      "Regions (
        region_id INTEGER PRIMARY KEY,
        region TEXT NOT NULL
      );"
    ),
    build_table(
      "Languages (
        lang_id INTEGER PRIMARY KEY,
        language TEXT NOT NULL
      );"
    ),
    build_table(
      "Also_Known_As (
        tconst TEXT NOT NULL,
        ordering INTEGER NOT NULL,
        title TEXT,
        region_id INTEGER,
        lang_id INTEGER,
        isOriginalTitle INTEGER, -- 0=false, else true
        PRIMARY KEY (tconst, ordering),
        FOREIGN KEY (tconst) REFERENCES Media(tconst),
        FOREIGN KEY (region_id) REFERENCES Regions(region_id),
        FOREIGN KEY (lang_id) REFERENCES Languages(lang_id)
      );"
    ),
    build_table(
      "Types (
        type_id INTEGER PRIMARY KEY,
        type TEXT NOT NULL
      );"
    ),
    build_table(
      "Aka_Types (
        akt_id INTEGER PRIMARY KEY,
        tconst TEXT NOT NULL,
        ordering INTEGER NOT NULL,
        type_id INTEGER NOT NULL,
        -- Media table PK tconst has ref. integrity through AKA Table
        FOREIGN KEY (tconst, ordering) REFERENCES Also_Known_As(tconst, ordering),
        FOREIGN KEY (type_id) REFERENCES Types(type_id),
        CONSTRAINT unique_akat_tc_ord_tid UNIQUE (tconst, ordering, type_id)
      );"
    ),
    build_table(
      "Attributes (
        att_id INTEGER PRIMARY KEY,
        att_name TEXT NOT NULL
      );"
    ),
    build_table(
      "Aka_Attributes (
        aka_id INTEGER PRIMARY KEY,
        tconst TEXT NOT NULL,
        ordering INTEGER NOT NULL,
        att_id INTEGER NOT NULL,
        -- Media table PK tconst has ref. integrity through AKA Table
        FOREIGN KEY (tconst, ordering) REFERENCES Also_Known_As(tconst, ordering),
        FOREIGN KEY (att_id) REFERENCES Attributes(att_id),
        CONSTRAINT unique_akaatt_tc_ord_aid UNIQUE (tconst, ordering, att_id)
      );"
    )
  )
  
  # Actually create the tables
  for (table_stmt in tables) {
    dbExecute(conn, table_stmt)
  }
}
```

## Reading Data Locally

Next we need to gather. Because of the size of the data set, we will take a small sample of it for the purposes of this practicum.

We'll set up some constants first.

```{r}
dir <- "./sample-data/"
ext <- ".sample.tsv"
names <- paste(dir, "name.basics", ext, sep="")
titles <- paste(dir, "title.basics", ext, sep="")
akas <- paste(dir, "title.akas", ext, sep="")
episodes <- paste(dir, "title.episode", ext, sep="")
principals <- paste(dir, "title.principals", ext, sep="")
directors <- paste(dir, "title.directors", ext, sep="")
writers <- paste(dir, "title.writers", ext, sep="")
ratings <- paste(dir, "title.ratings", ext, sep="")
```

### Titles

The number of titles we choose to include will determine the amount of data read in from other files. Therefore we will get this data first.

```{r, message=FALSE}
if (load) {
  df.titles <- read_tsv(titles)
  head(df.titles)
}
```

### AKAs

```{r, message=FALSE}
if (load) {
  df.akas <- read_tsv(akas)
  head(df.akas)
}
```

### Episodes

```{r, message=FALSE}
if (load) {
  df.eps <- read_tsv(episodes)
  head(df.eps)
}
```

### Ratings

```{r, message=FALSE}
if (load) {
  df.ratings <- read_tsv(ratings)
  head(df.ratings)
}
```

### Principles

```{r, message=FALSE}
if (load) {
  df.princ <- read_tsv(principals)
  head(df.princ)
}
```

### Crew

In pre-processing the crew data was split into separate sets for directors and writers.

#### Directors

```{r, message=FALSE}
if (load) {
  df.dirs <- read_tsv(directors)
  head(df.dirs)
}
```

#### Writers

```{r, message=FALSE}
if (load) {
  df.writers <- read_tsv(writers)
  head(df.writers)
}
```

### Names/People

```{r, message=FALSE}
if (load) {
  df.names <- read_tsv(names)
  head(df.names)
}
```

## Loading Data to Database

This section will take the data extracted previously, manipulate it as needed, and then add it to our database as needed.

To show our additions worked, we will show the top 5 rows that were just added to the current table. The following function provides the SQL to do this.

```{r}
headSql <- function(tableName) {
  paste("SELECT * FROM", tableName, "LIMIT 5;")
}
```


### "Formats" table

```{r}
loadFormats <- function(){
  # Grab the needed columns from the current dataframe
  types <- df.titles[,c(0,2)]
  # Since this is a categorical table, grab unique values
  format <- unique(types)
  # Create a new dataframe with unique values
  df_tt <- data.frame(format)
  # Remove titles with no format listed
  df_tt <- drop_na(df_tt)
  # Assign unique ids to the unique values - only works on initial loading of data
  df_tt <- tibble::rowid_to_column(df_tt, "format_id")
  colnames(df_tt) <- c("format_id", "format")
  # Transfer the new dataframe into the database table
  # Append is true because we want to maintain the relationships during table creation
  dbWriteTable(conn,"Formats",df_tt, append=TRUE)
  dbGetQuery(conn, headSql("Formats"))
  return(df_tt)
}
```

### "Media" table

```{r}
loadMedia <- function(df_tt){
  # Grab the needed columns from the current dataframe
  df_media <- df.titles[,-c(0,9)]
  # Add a column for the FK id
  df_media <- add_column(df_media, format_id = NA, .after = 1)
  
  # Get a list of the categories available
  types <- df_tt[,c(0,2)]
  # For each category, assign the FK id
  for(term in types) {
    search1 <- df_tt$format == term
    # Get the FK id for the category
    id <- df_tt[which(search1),][["format_id"]]
    search2 <- df_media$titleType == term
    # Replace default 0 value with the FK id
    df_media$format_id[search2] <- id
  }
  
  # Since the original default value was set to NA, when it is transferred 
  # to the database, those values will change to NULL. 
  # No additional checks are needed
  
  # Drop the category column, only the FK id column is needed
  df_media$titleType <- NULL
  
  # Transfer the new dataframe into the database table
  # Append is true because we want to maintain the relationships during table creation
  dbWriteTable(conn,"Media",df_media, append=TRUE)
  dbGetQuery(conn, headSql("Media"))
}
```

### "Genres" and "Media_Genres" tables

```{r}
loadGenres <- function(){
  # Grab the needed columns from the current dataframe
  df_mg <- df.titles[,c(1,9)]
  
  # Separate multi-valued attributes
  df_mg <- df_mg %>% 
    mutate(genres = strsplit(genres, ",")) %>%
    unnest(genres)
  # Assign unique ids to the unique values - only works on initial loading of data
  df_mg <- tibble::rowid_to_column(df_mg, "mg_id")
  # Remove titles with no genre listed
  df_mg <- drop_na(df_mg)
  # Add a column for the FK id
  df_mg <- add_column(df_mg, genre_id = NA, .after = 2)
  
  # Since genres is a categorical table, grab unique values
  df_genres <- unique(df_mg["genres"])
  # Assign unique ids to the unique values - only works on initial loading of data
  df_genres <- tibble::rowid_to_column(df_genres, "genre_id")
  # Rename column names to match table
  colnames(df_genres) <- c("genre_id", "genre")
  
  # Convert the genre into the appropriate FK id
  # Get a vector list of the categories available
  types <- df_genres[["genre"]]
  # For each category, assign the FK id
  for(term in types) {
    search1 <- df_genres$genre == term
    # Get the FK id for the category
    id <- df_genres[which(search1),][["genre_id"]]
    search2 <- df_mg$genres == term
    # Replace default 0 value with the FK id
    df_mg$genre_id[search2] <- id
  }
  
  # Since the original default value was set to NA, when it is transferred 
  # to the database, those values will change to NULL. 
  # No additional checks are needed
  
  # Drop the category column, only the FK id column is needed
  df_mg$genres <- NULL
  
  # Transfer the new dataframes into the database tables
  # Append is true because we want to maintain the relationships during table creation
  dbWriteTable(conn,"Genres",df_genres, append=TRUE)
  dbWriteTable(conn,"Media_Genres",df_mg, append=TRUE)
  dbGetQuery(conn, headSql("Genres"))
  dbGetQuery(conn, headSql("Media_Genres"))
}
```

### "Episodes" table

```{r}
loadEpisodes <- function(){
  # Transfer the new dataframes into the database tables
  # Append is true because we want to maintain the relationships during table creation
  dbWriteTable(conn,"Episodes",df.eps, append=TRUE)
  dbGetQuery(conn, headSql("Episodes"))
}
```

###"Ratings" table

```{r}
loadRatings <- function(){
  # Transfer the new dataframes into the database tables
  # Append is true because we want to maintain the relationships during table creation
  dbWriteTable(conn,"Ratings",df.ratings, append=TRUE)
  dbGetQuery(conn, headSql("Ratings"))
}
```

### "People" table

```{r}
loadPeople <- function(){
  # Grab single valued columns
  df.namesTable <- df.names[, c(1,2,3,4)]
  
  # Add age and numMovie columns
  # Leave NA for now, will use update statements to change later
  df.namesTable$age <- as.integer(NA)
  df.namesTable$numMovies <- as.integer(NA)
  
  # Add to table
  # Append is true because we want to maintain the relationships during table creation
  dbWriteTable(conn, "People", df.namesTable, append=TRUE)
  dbGetQuery(conn, headSql("People"))
}
```

### "Known_For_Titles" table

```{r}
loadKFT <- function(){
  # Extract the titles that people are known for and flatten it out
  # so only one name and one title appear in the same record.
  df.knownForTitles <- drop_na(df.names[c("nconst", "knownForTitles")]) %>%
      mutate(knownForTitles = strsplit(as.character(knownForTitles), ",")) %>%
      unnest(knownForTitles)
  
  # Add PK column based on row number
  df.knownForTitles <- tibble::rowid_to_column(df.knownForTitles, "kt_id")
  
  # Put in database
  # Append is true because we want to maintain the relationships during table creation
  dbWriteTable(conn, "Known_For_Titles", df.knownForTitles, overwrite=TRUE)
  dbGetQuery(conn, headSql("Known_For_Titles"))
}
```

### "Professions" and "Primary_Profession" table

```{r}
loadProfessions <- function(){
  # Extract the titles that people are known for and flatten it out
  # so only one name and one title appear in the same record.
  df.primaryProfessions <- drop_na(df.names[c("nconst", "primaryProfession")]) %>%
      mutate(primaryProfession = strsplit(as.character(primaryProfession), ",")) %>%
      unnest(primaryProfession)
  
  # Extract unique professions
  professions <- unique(df.primaryProfessions$primaryProfession)
  df.professions <- data.frame(professions)
  
  # Add PK columns
  df.professions <- tibble::rowid_to_column(df.professions, "prof_id")
  df.primaryProfessions <- tibble::rowid_to_column(df.primaryProfessions, "pp_id")
  
  # Rename columns as needed
  df.professions <- df.professions %>%
    rename(prof_title = professions)
  df.primaryProfessions <- df.primaryProfessions %>%
    rename(prof_id = primaryProfession)
  
  # Replace profession strings with an id
  profList <- df.professions$prof_title
  for (i in 1:length(profList)) {
    prof <- profList[i]
    df.primaryProfessions$prof_id[which(
      df.primaryProfessions$prof_id == prof
    )] <- i
  }
  df.primaryProfessions <- transform(
    df.primaryProfessions, prof_id = as.integer(prof_id))
  
  # Add to database
  # Append is true because we want to maintain the relationships during table creation
  dbWriteTable(conn, "Professions", df.professions, append=TRUE)
  dbWriteTable(conn, "Primary_Profession", df.primaryProfessions, append=TRUE)
  dbGetQuery(conn, headSql("Professions"))
  dbGetQuery(conn, headSql("Primary_Profession"))
}
```

### "Categories" table

```{r}
loadCategories <- function(){
  # Grab the needed columns from the current dataframe
  types <- df.princ[,c(0,4)]
  # Since this is a categorical table, grab unique values
  format <- unique(types)
  # Create a new dataframe with unique values
  df_cat <- data.frame(format)
  # Assign unique ids to the unique values - only works on initial loading of data
  df_cat <- tibble::rowid_to_column(df_cat, "category_id")
  # Transfer the new dataframe into the database table
  # Append is true because we want to maintain the relationships during table creation
  dbWriteTable(conn,"Categories",df_cat, append=TRUE)
  dbGetQuery(conn, headSql("Categories"))
  return(df_cat)
}
```

### "Crew" table

```{r}
loadCrew <- function(df_cat){
  # Grab the needed columns from the current dataframe
  df_crew <- df.princ[,-c(0,6)]
  # Add a column for the FK id
  df_crew <- add_column(df_crew, category_id = NA, .after = 3)
  
  # Get a list of the categories available
  types <- df_cat[,c(0,2)]
  # For each category, assign the FK id
  for(term in types) {
    search1 <- df_cat$category == term
    # Get the FK id for the category
    id <- df_cat[which(search1),][["category_id"]]
    search2 <- df_crew$category == term
    # Replace default 0 value with the FK id
    df_crew$category_id[search2] <- id
  }
  
  # Since the original default value was set to NA, when it is transferred 
  # to the database, those values will change to NULL. 
  # No additional checks are needed
  
  # Drop the category column, only the FK id column is needed
  df_crew$category <- NULL
  
  # Transfer the new dataframe into the database table
  # Append is true because we want to maintain the relationships during table creation
  dbWriteTable(conn,"Crew",df_crew, append=TRUE)
  dbGetQuery(conn, headSql("Crew"))
}
```

### "Cast" and "Characters" tables

```{r}
loadCast <- function(){
  # Grab the needed columns from the current dataframe
  df_cast <- df.princ[,-c(3,4,5)]
  # Remove titles with no genre listed
  df_cast <- drop_na(df_cast)
  
  # Remove brackets from data
  df_cast <- df_cast %>% 
    mutate(characters = str_replace(characters,"\\[",""))
  df_cast <- df_cast %>% 
    mutate(characters = str_replace(characters,"\\]",""))
  
  # Separate multi-valued attributes
  df_cast <- df_cast %>% 
    mutate(characters = strsplit(characters, "\",\"")) %>%
    unnest(characters)
  
  # Remove quotes from data
  df_cast <- df_cast %>% 
    mutate(characters = str_replace(characters,"\"",""))
  df_cast <- df_cast %>% 
    mutate(characters = str_replace(characters,"\"",""))
  
  # Assign unique ids to the unique values - only works on initial loading of data
  df_cast <- tibble::rowid_to_column(df_cast, "cast_id")
  # Add a column for the FK id
  df_cast <- add_column(df_cast, character_id = NA, .after = 3)
  
  # Since genres is a categorical table, grab unique values
  df_characters <- unique(df_cast["characters"])
  # Assign unique ids to the unique values - only works on initial loading of data
  df_characters <- tibble::rowid_to_column(df_characters, "character_id")
  # Rename column names to match table
  colnames(df_characters) <- c("character_id", "character")
  
  # Convert the genre into the appropriate FK id
  # Get a vector list of the categories available
  types <- df_characters[["character"]]
  # For each category, assign the FK id
  for(term in types) {
    search1 <- df_characters$character == term
    # Get the FK id for the category
    id <- df_characters[which(search1),][["character_id"]]
    search2 <- df_cast$characters == term
    # Replace default 0 value with the FK id
    df_cast$character_id[search2] <- id
  }
  
  # Since the original default value was set to NA, when it is transferred 
  # to the database, those values will change to NULL. 
  # No additional checks are needed
  
  # Drop the category column, only the FK id column is needed
  df_cast$characters <- NULL
  
  # Transfer the new dataframes into the database tables
  # Append is true because we want to maintain the relationships during table creation
  dbWriteTable(conn,"Characters",df_characters, append=TRUE)
  dbWriteTable(conn,"Cast",df_cast, append=TRUE)
  dbGetQuery(conn, headSql("Characters"))
  dbGetQuery(conn, headSql("Cast"))
}
```

### "Media_Directors" table

```{r}
loadDirectors <- function(){
  # Assign unique ids to the unique values - only works on initial loading of data
  df_direct <- tibble::rowid_to_column(df.dirs, "md_id")
  # Rename the columns to match the database table
  colnames(df_direct) <- c("md_id", "tconst", "nconst")
  # Transfer the new dataframes into the database tables
  # Append is true because we want to maintain the relationships during table creation
  dbWriteTable(conn,"Media_Directors",df_direct, append=TRUE)
  dbGetQuery(conn, headSql("Media_Directors"))
}
```

### "Media_Writers" table

```{r}
loadWriters <- function(){
  # Assign unique ids to the unique values - only works on initial loading of data
  df_write <- tibble::rowid_to_column(df.writers, "mw_id")
  # Rename the columns to match the database table
  colnames(df_write) <- c("mw_id", "tconst", "nconst")
  # Transfer the new dataframes into the database tables
  # Append is true because we want to maintain the relationships during table creation
  dbWriteTable(conn,"Media_Writers",df_write, append=TRUE)
  dbGetQuery(conn, headSql("Media_Writers"))
}
```

### "Regions" table

```{r}
loadRegions <- function(){
  # Grab the needed columns from the current dataframe
  types <- df.akas[,c(0,4)]
  # Since this is a categorical table, grab unique values
  format <- unique(types)
  # Create a new dataframe with unique values
  df_reg <- data.frame(format)
  # Remove titles with no genre listed
  df_reg <- drop_na(df_reg)
  # Assign unique ids to the unique values - only works on initial loading of data
  df_reg <- tibble::rowid_to_column(df_reg, "region_id")
  # Transfer the new dataframe into the database table
  # Append is true because we want to maintain the relationships during table creation
  dbWriteTable(conn,"Regions",df_reg, append=TRUE)
  dbGetQuery(conn, headSql("Regions"))
  return(df_reg)
}
```

### "Languages" table
```{r}
loadLangs <- function(){
  # Grab the needed columns from the current dataframe
  types <- df.akas[,c(0,5)]
  # Since this is a categorical table, grab unique values
  format <- unique(types)
  # Create a new dataframe with unique values
  df_lang <- data.frame(format)
  # Remove titles with no genre listed
  df_lang <- drop_na(df_lang)
  # Assign unique ids to the unique values - only works on initial loading of data
  df_lang <- tibble::rowid_to_column(df_lang, "lang_id")
  # Transfer the new dataframe into the database table
  # Append is true because we want to maintain the relationships during table creation
  dbWriteTable(conn,"Languages",df_lang, append=TRUE)
  dbGetQuery(conn, headSql("Languages"))
  return(df_lang)
}
```

### "Also_Known_As" table

```{r}
loadAKA <- function(df_reg, df_lang){
  # Grab the needed columns from the current dataframe
  df_aka <- df.akas[,-c(6,7)]
  # Rename the columns to match the database table
  colnames(df_aka) <- c("tconst", "ordering", "title", "region",
                        "language", "isOriginalTitle")
  
  # Add a column for the FK id
  df_aka <- add_column(df_aka, region_id = NA, .after = 3)
  
  # Get a list of the categories available
  types <- df_reg[,c(0,2)]
  # For each category, assign the FK id
  for(term in types) {
    search1 <- df_reg$region == term
    # Get the FK id for the category
    id <- df_reg[which(search1),][["region_id"]]
    search2 <- df_aka$region == term
    # Replace default 0 value with the FK id
    df_aka$region_id[search2] <- id
  }
  
  # Add a column for the FK id
  df_aka <- add_column(df_aka, lang_id = NA, .after = 5)
  
  # Get a list of the categories available
  types <- df_lang[,c(0,2)]
  # For each category, assign the FK id
  for(term in types) {
    search1 <- df_lang$language == term
    # Get the FK id for the category
    id <- df_lang[which(search1),][["lang_id"]]
    search2 <- df_aka$language == term
    # Replace default 0 value with the FK id
    df_aka$lang_id[search2] <- id
  }
  
  # Since the original default values were set to NA, when it is transferred 
  # to the database, those values will change to NULL. 
  # No additional checks are needed
  
  # Drop the category columns, only the FK id columns are needed
  df_aka$region <- NULL
  df_aka$language <- NULL
  
  # Transfer the new dataframe into the database table
  # Append is true because we want to maintain the relationships during table creation
  dbWriteTable(conn,"Also_Known_As",df_aka, append=TRUE)
  dbGetQuery(conn, headSql("Also_Known_As"))
}
```

### "Types" and "Aka_Types" tables

```{r, message=FALSE}
loadTypes <- function(){
  # Get aka types
  df.akaTypes <- drop_na(df.akas[c("titleId", "ordering", "types")])
  
  # Extract unique types
  types <- unique(df.akas$types)
  df.types <- data.frame(types)
  df.types <- drop_na(df.types)
  
  # Add PK columns
  df.types <- tibble::rowid_to_column(df.types, "type_id")
  df.akaTypes <- tibble::rowid_to_column(df.akaTypes, "akt_id")
  
  # Rename and change type of columns as needed
  df.types <- df.types %>%
    rename(type = types)
  df.types <- transform(
    df.types, type = as.character(type))
  
  colnames(df.akaTypes) <- c("akt_id", "tconst", "ordering", "type_id")
  
  # Replace type strings with an id
  typeList <- df.types$type
  for (i in 1:length(typeList)) {
    type <- typeList[i]
    df.akaTypes$type_id[which(
      df.akaTypes$type_id == type
    )] <- i
  }
  df.akaTypes <- transform(
    df.akaTypes,
    type_id = as.integer(type_id),
    ordering = as.integer(ordering)
  )
  
  # Add to database
  # Append is true because we want to maintain the relationships during table creation
  dbWriteTable(conn, "Types", df.types, append=TRUE)
  dbWriteTable(conn, "Aka_Types", df.akaTypes, append=TRUE)
  dbGetQuery(conn, headSql("Types"))
  dbGetQuery(conn, headSql("Aka_Types"))
}
```

### "Attributes" and "Aka_Attributes" tables

```{r, message=FALSE}
loadAttributes <- function(){
  # Get aka attributes
  df.akaAtts <- drop_na(df.akas[c("titleId", "ordering", "attributes")])
  
  # Extract unique attributes
  attributes <- unique(df.akas$attributes)
  df.attributes <- data.frame(attributes)
  df.attributes <- drop_na(df.attributes)
  
  # Add PK columns
  df.attributes <- tibble::rowid_to_column(df.attributes, "att_id")
  df.akaAtts <- tibble::rowid_to_column(df.akaAtts, "aka_id")
  
  # Rename and change type of columns as needed
  df.attributes <- df.attributes %>%
    rename(att_name = attributes)
  df.attributes <- transform(
    df.attributes, att_name = as.character(att_name))
  
  colnames(df.akaAtts) <- c("aka_id", "tconst", "ordering", "att_id")
  
  # Replace type strings with an id
  attList <- df.attributes$att_name
  for (i in 1:length(attList)) {
    att <- attList[i]
    df.akaAtts$att_id[which(
      df.akaAtts$att_id == att
    )] <- i
  }
  
  df.akaAtts <- transform(
    df.akaAtts,
    # introduces some NAs by coercion, this is OK, there are many NAs
    att_id = as.integer(att_id),
    ordering = as.integer(ordering)
  )
  
  # Add to database
  # Append is true because we want to maintain the relationships during table creation
  dbWriteTable(conn, "Attributes", df.attributes, append=TRUE)
  dbWriteTable(conn, "Aka_Attributes", df.akaAtts, append=TRUE)
  dbGetQuery(conn, headSql("Attributes"))
  dbGetQuery(conn, headSql("Aka_Attributes"))
}
```


### Load the data into the tables
```{r}
if (load) {
  df_cat1 <- loadFormats()
  loadMedia(df_cat1)
  loadGenres()
  loadEpisodes()
  loadRatings()
  loadPeople()
  loadKFT()
  loadProfessions()
  df_cat1 <- loadCategories()
  #loadCrew(df_cat1)
  #loadCast()
  loadDirectors()
  loadWriters()
  df_cat1 <- loadRegions()
  df_cat2 <- loadLangs()
  loadAKA(df_cat1, df_cat2)
  loadTypes()
  loadAttributes()
}
```


## Update Tables

```{r}

```


## Triggers

```{r}

```


## View

```{sql connection=conn}
DROP VIEW IF EXISTS actors;
```

```{sql connection=conn}
CREATE VIEW actors
(name, age, alive, numMovies)
AS
SELECT 
  primaryName, 
  age, 
  CASE WHEN deathYear is NULL THEN "No" ELSE "Yes" END, 
  titles 
FROM people
JOIN
  (SELECT counts.nconst, prof_id, titles 
  FROM Primary_Profession pp 
  JOIN 
    (SELECT nconst, COUNT(nconst) AS titles 
    FROM Known_For_Titles 
    GROUP BY nconst) counts 
  ON counts.nconst=pp.nconst
  WHERE prof_id IN (2,4)) limited
ON people.nconst=limited.nconst;
```

```{sql connection=conn}
SELECT * FROM actors LIMIT 5;
```

## Queries

###Task 8: Write a query that finds the number of seasons for each TV series. Using the results of the query create a histogram (frequency plot) with proper axis labels and title.

```{r}

```

```{sql connection=conn}
SELECT primaryTitle, parentTconst, MAX(seasonNumber) AS numSeasons
FROM Episodes 
JOIN 
  (SELECT tconst, primaryTitle 
  FROM Media 
  WHERE format_id=2) series 
ON series.tconst=Episodes.parentTconst
GROUP BY parentTconst;
```


###Task 11: Write a query to retrieve the names and ages of all actors who appeared in more than two movies (but not TV Movies) which an above average rating. Show the results of the query in your notebook. Do not hard code the average rating.

```{r}
average <- dbGetQuery(conn, "SELECT AVG(averageRating) AS value FROM Ratings")[["value"]]
average
```

```{sql connection=conn}
SELECT * FROM Formats;
```

```{sql connection=conn}
SELECT People.nconst, primaryName, age 
FROM People 
JOIN
  (SELECT nconst
  FROM Primary_Profession pp 
  WHERE prof_id IN (2,4)) actors
ON people.nconst=actors.nconst LIMIT 5;
```

```{sql connection=conn}
SELECT Media.tconst, averageRating
FROM Media 
JOIN
  (SELECT tconst, averageRating
  FROM Ratings 
  WHERE averageRating > ?average) above
ON above.tconst=Media.tconst
WHERE format_id = 1;
```

###Task 12: Write a query that finds an actor by name (pick a name). Measure the execution time of the query. Then create an index that would improve the performance of the query and then run and measure it again. Show the difference in a bar chart and comment on why that's the case.

```{r}

```

```{sql connection=conn}
SELECT * FROM actors LIMIT 5;
```

